\section{Illustrative Examples}

To demonstrate the functionalities of the \href{https://github.com/MatlabCompat/MatlabCompat.jl}{MatlabCompat.jl} library we have created \textit{\textit{\href{https://github.com/MatlabCompat/MatlabCompat.jl/blob/master/test/janus.m}{janus.m}}} - a minimalistic image analysis program in Matlab/Octave (see \textbf{supplementary video S1}). First, the program is reading a remote image - a micrograph of cultured cells, stained with a nuclear stain (A549 human lung carcinoma cells stained with Hoechst 33342 dye from a dataset used it \cite{22787215}). Next the program is using Otsu \cite{otsu1975threshold} thresholding to segment (separate) bright foreground pixels from dark background pixels. Finally, the program counts connected components in the foreground pixels, which in an idealistic case is equal to the number of cellular nuclei in the picture. \textbf{Fig. 3A} shows MATLAB\textsuperscript{TM} R2014a output upon invoking \textit{janus.m}. Code listing for \textit{janus.m} together with commentaries (separated by the symbol "\%" in Matlab/Octave) is provided bellow:\\



\verb|tic()|

\verb|% read the remote image|

\verb|img = imread('http://matlabcompat.github.io/img/example.tif');|

\verb|% compute graysacale threshold using Otsu algorithm|

\verb|threshold = graythresh(img);|

\verb|% create a binary image based on the grayscale image|

\verb|imgbw = im2bw(img, threshold);|

\verb|% display the resulting binary image|

\verb|imshow(imgbw);|

\verb|% lable each connected object in the image|

\verb|labeledbw = bwlabel(imgbw, 4);|

\verb|% count cells|

\verb|numberOfCells = max(reshape(labeledbw, 1,numel(labeledbw)));|

\verb|% display the number of objects|

\verb|disp(strcat('number of objects:', num2str(numberOfCells)));|

\verb|imshow(label2rgb(labeledbw,'jet',[0 0 0],'shuffle'));|

\verb|toc()|\\



To ensure that the same program can be understood by Julia using \href{https://github.com/MatlabCompat/MatlabCompat.jl}{MatlabCompat.jl} package either a parsing step or minimal changes to the code are necessary. These changes include basic syntax adaptation like substitution of single quotes symbol (') by double quotes ("), substitution of commentary symbol percent (\%) by hash (\#) and addition of the lines initiating packages used including at least MatlabCompat. This is basic parsing is provided by the function \textit{"rosetta()"}, like in the example below (identical to the example mentioned above):\\



\verb|using MatlabCompat|

\verb|rosetta(janus.m, janus.jl)|\\



In case of \textit{\href{https://github.com/MatlabCompat/MatlabCompat.jl/blob/master/test/janus.m}{janus.m}} the function \textit{"rosetta()"} generates immediately Julia interpretable code (see \textbf{Fig. 3B} for output in Julia 0.3.10):\\



\verb|using MatlabCompat|

\verb|importall MatlabCompat.ImageTools|

\verb|importall MatlabCompat.MathTools|

\verb|tic()|

\verb|# read the remote image|

\verb|img = imread("http://matlabcompat.github.io/img/example.tif");|

\verb|# compute graysacale threshold using Otsu algorithm|

\verb|threshold = graythresh(img);|

\verb|# create a binary image based on the grayscale image|

\verb|imgbw = im2bw(img, threshold);|

\verb|# display the resulting binary image|

\verb|imshow(imgbw);|

\verb|# lable each connected object in the image|

\verb|labeledbw = bwlabel(imgbw, 4);|

\verb|# count cells|

\verb|numberOfCells = max(reshape(labeledbw, 1,numel(labeledbw)));|

\verb|disp(strcat("number of objects:", num2str(numberOfCells)));|

\verb|# display the number of objects|

\verb|imshow(label2rgb(labeledbw,"jet",[1 1 1],"shuffle"));|

\verb|toc()|\\


To show how similarly simplistic microprogram, which can be readily automatically translated from MATLAB/Octave to Julia , can be used to address a research question in a different scientific filed we have used another \href{http://www.broadinstitute.org/bbbc/BBBC010/}{published dataset} of an image-based screen performed for roundworm \textit{Caenorhabditis elegans} model organism \cite{Moy_2009}. The authors of that screen aimed to look for novel small compounds inhibiting \textit{Enterococcus faecalis} bacterial infection. Their readout was fluorescence signal of SYTOX Orange dye staining specifically dead worms, while living worms remained impermeable to that dye - thus emitting little or no fluorescent signal. Death of the experimental \textit{C. elegans} worms in that study was attributed to \textit{E. faecalis} bacterial infection (\textbf{Fig. 3C}). For illustration purposes, of how a simple MATLAB/Octave or Julia programs can measure such a readout we have first created the \textit{\href{http://}{janus2.m}} MATLAB/Octave program counting foreground pixels from a fluorescence micrograph (\textbf{Fig. 3D}): 

\verb|% read the remote images|

\verb|imgTreated = imread('http://matlabcompat.github.io/img/|

\verb|example_Treated.tif');|

\verb|imgNonTreated = imread('http://matlabcompat.github.io/|

\verb|img/example_NonTreated.tif');|

\verb|% compute grayscale threshold of the non treated image|

\verb|% using Otsu algorithm|

\verb|threshold = graythresh(imgNonTreated);|

\verb|% create a binary image based on the grayscale images|

\verb|imgBWTreated = im2bw(imgTreated, threshold);|

\verb|imgBWNonTreated = im2bw(imgNonTreated, threshold);|

\verb|% count foreground pixels for each image and output it|

\verb|% in console|

\verb|foregroundPixelCountTreated = sum(reshape(imgBWTreated,|

\verb| 1,numel(imgBWTreated)));|

\verb|foregroundPixelCountNonTreated =  sum(reshape(|

\verb|imgBWNonTreated, 1,numel(imgBWNonTreated)));|

\verb|display(strcat('Foreground pixel count for Treated |

\verb|specimen: ',num2str(foregroundPixelCountTreated)))|

\verb|display(strcat('Foreground pixel count for Non-Treated|

\verb|specimen: ',num2str(foregroundPixelCountNonTreated)))|\\


Next, similarly to the previous example, we invoked \textit{"rosetta()"} function on the \textit{\href{http://}{janus2.m}} to automatically translate it to Julia language, yielding \textit{\href{http://}{janus2.jl}}:


\verb|using MatlabCompat|

\verb|importall MatlabCompat.ImageTools|

\verb|importall MatlabCompat.MathTools|

\verb|# read the remote images|

\verb|imgTreated = imread("http://matlabcompat.github.io/|

\verb|img/example_Treated.tif");|

\verb|imgNonTreated = imread("http://matlabcompat.github.io/|

\verb|img/example_NonTreated.tif");|

\verb|# compute grayscale threshold of the non treated image|

\verb|# using Otsu algorithm|

\verb|threshold = graythresh(imgNonTreated);|

\verb|# create a binary image based on the grayscale images|

\verb|imgBWTreated = im2bw(imgTreated, threshold);|

\verb|imgBWNonTreated = im2bw(imgNonTreated, threshold);|

\verb|# count foreground pixels for each image and output it|

\verb|# in console|

\verb|foregroundPixelCountTreated = sum(reshape(imgBWTreated,|

\verb| 1,numel(imgBWTreated)));|

\verb|foregroundPixelCountNonTreated =  sum(reshape(imgBWNonTreated,|

\verb| 1,numel(imgBWNonTreated)));|

\verb|display(strcat("Foreground pixel count for Treated|

\verb|specimen: ",num2str(foregroundPixelCountTreated)))|

\verb|display(strcat("Foreground pixel count for Non-Treated|

\verb|specimen: ",num2str(foregroundPixelCountNonTreated)))|\\


Both programs gave identical fluorescence micrographs quantification results shown in \textbf{Fig. 3D}, suggesting that, as expected by the design of this published study \cite{Moy_2009}, the worm death was lower in the case of treatment with control inhibitor of \textit{E. faecalis} bacterial infection related worm death compared to the mock control. Altogether, these examples suggest, that in the specific case of MATLAB/Octave to Julia conversion, results, readily applicable to complex research questions, can be achieved rather quickly. While, as discussed before, this might not work equally well in every case, this conversion may give better results with further development of the \href{https://github.com/MatlabCompat/MatlabCompat.jl}{MatlabCompat.jl} by its community.


Arguably, one of the most attractive features Julia may provide for the research software is the potential for performance optimization. To challenge the results \href{http://julialang.org/benchmarks/}{published} on the official website of the language we have executed the published standard algorithms performance tests in slightly different environment for a new version of Julia (0.3.10) and two different versions of MATLAB\textsuperscript{TM} (R2014a and R2015a). The tests include Fibonacci recursion (fib), parse integers (parse\_int), Mandelbrot set: complex arithmetic and comprehensions (mandel), numeric vector sorting (quicksort), slow $\pi$ series (pi\_sum), random matrix statistics (rand\_mat\_stat), large random number generation and multiplication of random matrices (rand\_mat\_mul).

The test environment in the published results was: Intel速 Xeon速 CPU E7-8850 2.00GHz CPU with 1TB of 1067MHz DDR3 RAM, running Linux in single core execution. In our case the test environment was Intel速 Core\textsuperscript{TM} i7-4790K CPU E7-8850 4.00GHz (8 CPUs) CPU with 16GB of 1600MHz DDR3 RAM, running Windows 8.1 Pro 64-bit (6.3, Build 9600) without restriction to a single core. Results (\textbf{Tab. 1}) suggest small differences in execution time. The differences are most pronounced for Julia's most efficient Fibonacci Recursion (fib). This can be attributed to the difference in test settings and the version of the language. Despite the small differences, we were able to faithfully reproduce the pattern and the orders of magnitudes of the published results. Surprisingly MATLAB\textsuperscript{TM} R2015a performed slightly slower than R2014a in our test environment.

\begin{table} 
    \begin{tabular}{cccc}
    \hline
        \textbf{Algorithm} & \textbf{Reported 2014a\textsuperscript{1}} & \textbf{Tested  2014a\textsuperscript{2}} & \textbf{Tested 2015a\textsuperscript{2}}\\
        \hline
        fib & 1989.78 & 1032.79 & 1101.73 \\ 
        \hline
        parse\_int & 971.90 & 350.81 & 408.64 \\ 
        \hline
        mandel & 46.17 & 47.69 & 47.62 \\ 
        \hline
        quicksort & 69.07 & 47.28 & 57.80 \\ 
        \hline
        pi\_sum & 1.28 & 1.09 & 1.10 \\ 
        \hline
        rand\_mat\_stat & 5.64 & 1.41 & 3.91 \\ 
        \hline
        rand\_mat\_mul & 1.03 & 2.84 & 2.86 \\
        \hline
    \end{tabular}
    \caption{\textbf{Julia to MATLAB Performance Tests Comparison Measured by Specific Algorithm Execution Time in our Test Environment}\textsuperscript{3}. The Specific Algorithms included Fibonacci recursion (fib), parse an int from a string (parse\_int), Mandelbrot set (mandel), Quicksort sorting algorithm (quicksort), Slow $\pi$-series (pi\_sum), Random matrix statistics (rand\_mat\_stat), Random number matrix generation multiplication (rand\_mat\_mul). Source code for the performance tests were taken from the Julia community as is. The respective implementations are available on Github for \href{https://github.com/JuliaLang/julia/blob/master/test/perf/micro/perf.jl}{Julia} and \href{https://github.com/JuliaLang/julia/blob/master/test/perf/micro/perf.m}{MATLAB/Octave}. \textsuperscript{1}Performance of MATLAB\textsuperscript{TM} (measured as execution time in the test environment specified) folds performance of Julia 0.3.7. \textsuperscript{2}Performance of MATLAB\textsuperscript{TM} folds performance of Julia 0.3.10. \textsuperscript{3}Intel速 Core\textsuperscript{TM} i7-4790K CPU E7-8850 4.00GHz (8 CPUs) CPU with 16GB of 1600MHz DDR3 RAM, running Windows 8.1 Pro 64-bit (6.3, Build 9600).}
\end{table}
  
We have next compared the execution times of our model image analysis program identical to \textit{"janus.m"} and \textit{"janus.jl"}. Median results of 1000 iterations in the test environment mentioned above suggested that MATLAB\textsuperscript{TM} R2014a performed 0.29 folds of Julia 0.3.10 and R2015a 0.30 folds of Julia 0.3.10. We concluded that performance of Julia and MATLAB\textsuperscript{TM} was comparable for this specific case. This may be caused by the fact that MATLAB\textsuperscript{TM} is highly optimized for vector operations used in this program.