\section{Motivation and significance}
\label{} 

\textbf{Introduce the scientific background and the motivation for developing the software.}

Conversion of the language of mathematics into the language of machine instructions, often defined as scientific and technical computing has been pioneered by the project FORTRAN \cite{Backus:1957:FAC}. Lergely building upon the success of its predecessor, Matlab/Octave is a scientific and technical computing programming language widely used through academic community. MATLAB\textsuperscript{TM} has been introduced in late 1970s as an open source programming language for technical computing and later commercialized by Mathworks Inc. Being product of Mathworks Inc., MATLAB\textsuperscript{TM} became popular among primarily engineers providing a flexible solutions for linear algebra, differential equations and a number of other problems engineers deal with on a daily basis \cite{moore2014matlab}. Itroduced in the end of 1980s its syntaxial sibling, GNU Octave, offered an open source alternative to the commercial product of Mathworks \cite{eaton1997gnu}.

Yet more than a tool for engineers, in the last three decades Matlab/Octave has become a \textit{de facto} default tool for developing scientific code, often used to solve problems it was not initially designed for. This, in turn, lead to an accumulation of prototype-like scientific software solutions unable to scale up to the promise and, thus, limiting the research work. To address this, in recent years a number of new programming languages and frameworks has been proposed including SciPy for Python \cite{jones2001open, Olivier_2002}, SCILAB \cite{Campbell_2009} and others. While some projects managed to successfully undertake the effort of rewriting their research software into one of these new frameworks \cite{17076895, 21349861}, for many this represents a major problem due to programming language differences.

A newly introduced open source dynamic high level technical programing language Julia aims at shattering the paradigm that high level prototype code has to be inherently inefficient \cite{bezanson2012julia, bezanson2014julia}.  Unlike other high level technical computing languages Julia offers abstract types, multiple dispatch of methods allowing reusing them dynamically, and other attractive features of a modern programming language. The source code written in Julia is executed using just-in-time (JIT) compilation based on its low level virtual machine (LLVM) may deliver performance matching performance of the iconic C language \cite{bezanson2012julia, bezanson2014julia}. Similar to \textit{Esperanto} of human languages, the syntax of Julia combines to a larger extend syntaxes of Matlab/Octave, Python, R, Perl and many other languages. However as close as it may seem to Matlab/Octave the code written in Matlab/Octave requires significant modifications to be understood by Julia.

\textbf{Explain why the software is important, and describe the exact (scientific) problem(s) it solves.}

Here we present MatlabCompat.jl a Julia language library aimed at simplifying the conversion of Matlab/Octave based research software to the novel Julia technical computing language by providing one way compatibility and MATLLAB/Octave to Julia code parsing modules. Using this library research groups with limited software development and computational resources can convert their Matlab/Octave code into Julia in a much easier way. This step may, in turn, improve usability, performance and prolong the lifetime of their software tools. The library is written in native Julia alone, available on Github under MIT open source license and is a part of Julia package repository.

\textbf{Indicate in what way the software has contributed (or how it will contribute in the future) to the process of scientific discovery; if available, this is to be supported by citing a research paper using the software.}

Matlab/Octave is widely used to develop complex research software, e.g. in the filed of image analysis \cite{Kiss_2014, 17076895}. Beyond image analysis MatlabCompat.jl spans already at its early stage of development a handful of functions including input/output, working with strings, linear algebra and others providing researchers with an easy-to-reach Matlab/Octave alternative. MatlabCompat.jl allows researchers to first easily convert their Matlab/Octave code into Julia and then enrich their code using Julia language. The added values of such code conversion efforts include using features like dynamic types, dataframes, as well as, allowing for additional performance optimization.

To illustrate how MatlabCompat.jl helps solving typical scientific problem of experimental results quantification we provide a typical use case from image analysis of micrographs obtained with an automated epifluorescent screening microscope (based on published data \cite{22787215}). E.g. to quantify fluorescent micrographs it is common to perform to separate background (dark) pixels from foreground (bright) pixels based on a threshold value of gray values determined either manually or using e.g. Otsu algorithm \cite{otsu1975threshold}. Together with other relevant functions, we have implemented the latter in Julia for a use case identical to a typical Matlab/Octave use case, described in detail below.

\textbf{Provide a description of the experimental setting (how does the user use the software?).}

Beyond image analysis, the range of the settings where MatlabCompat.jl can be used is as broad as scientific and technical computing themselves. It ranges from data acquisition, mining and analysis to modeling, simulation and prediction. It aims at providing added value to existing Matlab/Octave solutions by assisting in code conversion into Julia - a language with broader possibilities and prospective performance optimization.

To use MatlabCompat.jl upon \href{http://julialang.org/downloads/}{download} and installation of Julia distributive for the respective platform, the user has to add the package by invoking:

\verb|Pkg.add("MatlabCompat")|

and subsequently before calling specific functions from our package:

\verb|using MatlabCompat|

Basic parsing of Matlab/Octave code and its preconversion to Julia code is provided by the function \textit{rossetta()} using following syntax:

\verb|rossetta(path/to/your/Mfile.m, path/to/new/jlfile.jl)|

\textbf{Introduce related work in literature (cite or list algorithms used, other software etc.).}

One or two-way compatibility between different programming languages can be achieved either by manual or automatic translation, application program interface (API) based integration or, in cases of exceptional syntax similarity, by compatibility libraries. Multiple efforts on automatic translation of programming languages date to as early as 1960s \cite{Ledley_1962, irons1961syntax}. API based integrations exists between and for multiple programming languages and platforms, including e.g. Python \cite{autin2012upy}, MATLAB \cite{Bornstein_2008}. Julia languages has multiple APIs based integrations working and in development, including MATLAB.jl allowing calling MATLAB\textsuperscript{TM} using MATLAB\textsuperscript{TM} engine from Julia environment. Finally, compatibility libraries a typically written in the destination language and are a very common way to translate not just code but also principles and practices well established in one language to another language, like e.g. Matlab plotting library for Python \cite{Hunter_2007, barrett2005matplotlib}. While calling Matlab/Octave code from Julia is made possible by the MATLAB.jl package, there was no tool thus far assisting in conversion of Matlab/Octave 
code to Julia language. Here we propose to address this need using MatlabCompat.jl library package. We call for collaborative assistance on growing the functionality of the package, which can be accomplished by forking our source code through Github and pull-requesting your improvements later on (see details on contributing below). Additionally \href{http://matlabcompat.github.io/}{MatlabCompat.jl} is set up as an organization on Github aimed at uniting multiple researcher working on the package and creating a developers community.
    
    
    
    
    
    
    
    
    
    
    
    
    
    
    
    
  
  
  
  
  
  
  
  
  
  