\section{Software description}

\href{https://github.com/MatlabCompat/MatlabCompat.jl}{MatlabCompat.jl} is a Julia language package that helps converting your code in Matlab/Octave into Julia. Much like Julia language itself \href{https://github.com/MatlabCompat/MatlabCompat.jl}{MatlabCompat.jl} is fully Github based and a work in progress inviting everyone to collaborate. Following a Github workflow it is set up as a \href{https://github.com/MatlabCompat}{Github organization}, which is open for new members and the library source code is open for Github based pull requests. The longterm goal of the library is to allow effortless compatibility of Matlab/Octave research code to Julia.

A full (semantic) code conversion for high level programming languages is a highly non-trivial task. A more pragmatic and, thus, successful strategy would involve either partial translation/adaptation of the syntax of the source language to the syntax of the destination (object) language \cite{Ledley_1962} or usage of the library substituting (translating) the code of the source language "as is" in the environment of the destination language (\textbf{Fig. 1}). While first strategy is always available for a developer it's possibilities can be quickly exhausted. Therefore, the rest of the compatibility effort in \href{https://github.com/MatlabCompat/MatlabCompat.jl}{MatlabCompat.jl} is covered by the second strategies.

More often than not, certain code adaptation may become necessary or desired for several reasons. First,  both \href{https://github.com/MatlabCompat/MatlabCompat.jl}{MatlabCompat.jl} and Julia language itself are in an early stage. Therefore, \href{https://github.com/MatlabCompat/MatlabCompat.jl}{MatlabCompat.jl} might either not cover the necessary functionality of Matlab/Octave or this functionality might be non-compatible for syntaxial and other reasons. Second, pursuing exclusively strategy 2 is very challenging in development. Third, pursuing exclusively strategy 2 developers risk creating non-optimal \textit{"stub"} Julia code.

\href{https://github.com/MatlabCompat/MatlabCompat.jl}{MatlabCompat.jl} is helpful to simplify refactoring effort in transition from legacy Matlab/Octave code to new Julia code. The compatibility library aims to minimize naming conflicts between Matlab/Octave code reserved namespace and respective taken identifiers in the namespace of Julia. We refer to such property in newly converted code (strategy 1, see \textbf{Fig. 1}) as minimal target-invasive (Julia-invasive) compatibility.  Only in newly written or refactored (strategy 2, see \textbf{Fig. 1}) application code developers get fully exposed to utilization of all Julia features. Therefore developers are suggested to pursue hybrid strategy requiring code refactoring to make their Matlab/Octave work in Julia yet minimizing the effort needed to achieve it.

\subsection{Software Architecture}
\label{Architecture} 

Following the minimal Julia invasivity concept mentioned, \href{https://github.com/MatlabCompat/MatlabCompat.jl}{MatlabCompat.jl} has a well-structured architecture allowing minimization of namespace conflicts with Julia's \textit{Base} package (\textbf{Fig. 2}). A typical usage of a Julia package involves invoking \textit{"using"} command:

\verb|using MatlabCompat|

The namespace conflicts between Julia and Matlab/Octave are partially inevitable given the syntaxial closeness of Julia and Matlab/Octave. E.g. the function \textit{max()} in Matlab/Octave is closer to the function \textit{maximum()} in Julia, while \textit{max()} function name is taken in Julia. To address this type of issues Julia language avoids global declarations and uses local scopes. Typically for other Julia packages invoking using with a package would export all the functions of the package into the scope. Yet in case of \href{https://github.com/MatlabCompat/MatlabCompat.jl}{MatlabCompat.jl} to avoid namespace conflicts with \textit{"Base"} package, invoking the line above for MatlabCompat package would export into the scope only those functions names of which are not conflicting with \textit{"Base"} package. To substitute the respective function from the \textit{"Base"} package by a function from \href{https://github.com/MatlabCompat/MatlabCompat.jl}{MatlabCompat.jl} package with the same name (for Matlab/Octave compatibility reasons) a user would have to explicitly \textit{import} the respective module of \href{https://github.com/MatlabCompat/MatlabCompat.jl}{MatlabCompat.jl} with the same naming prefix of this module or a particular symbol within such module into the scope of the converted program. Nonetheless, most of these special cases are explicitly handled by symbol export within \href{https://github.com/MatlabCompat/MatlabCompat.jl}{MatlabCompat.jl}. For example to use max() function from \href{https://github.com/MatlabCompat/MatlabCompat.jl}{MatlabCompat.jl} package rather than from \textit{"Base"} package - write:\\

\verb|using MatlabCompat, MatlabCompat.MathTools|

\verb|max([1 2 3])|

\verb|3|\\


\href{https://github.com/MatlabCompat/MatlabCompat.jl}{MatlabCompat.jl} follows a rather simple modules organization convention. Namely, all the modules are named with function purposes based self-explanatory names in \textit{"CamelCase"} with capitalized first letter. Modules may contain submodules if necessary. All functions modules contain should be either aliases (wrappers) or re-implementations of Matlab/Octave functions named the same way. The only module representing an exception in this case is module \textit{"Support"}, which contains helper-functions, not having an analogue neither in Matlab/Octave nor in Julia languages, necessary for other functions in the package.

\subsection{Software Functionalities}

Currently the major functionalities of the \href{https://github.com/MatlabCompat/MatlabCompat.jl}{MatlabCompat.jl} include syntax aliases or reimplementations of Matlab/Octave functions organized in modules with self-explanatory naming: \textit{"StringTools"}, \textit{"MathTools"}, \textit{"ImageTools"}, \textit{"Support"} (see \textbf{Fig. 2} and \href{http://matlabcompat.github.io/help.html}{help}). Like the Julia language, \href{https://github.com/MatlabCompat/MatlabCompat.jl}{MatlabCompat.jl} is in active development and is actively receiving new contributions, including new modules. For an up-to-date overview of functions provided by different modules of the package please see the \href{http://matlabcompat.github.io/help.html}{help}. For pragmatic reasons the Matlab/Octave to Julia compatibility functions covered by \href{https://github.com/MatlabCompat/MatlabCompat.jl}{MatlabCompat.jl} at the current stage are strongly shifted towards image analysis and processing functions, including e.g. \textit{graythresh()} for Otsu thresholding \cite{otsu1975threshold}, \textit{bwlabel()} for labeling and subsequent counting of e.g. bright objects over dark background (discussed in detail below) and other functions.

\subsection{Sample code snippets analysis}

An example of functionality provided by the \href{https://github.com/MatlabCompat/MatlabCompat.jl}{MatlabCompat.jl} package standing out, but worth mentioning is covered by function \textit{rossetta()} (\textit{"Support" module}). This function is aimed to be used as the starting point for Matlab/Octave-to-Julia conversion of your code. It is aimed to parse m-files (first argument) making basic syntax changes required for the code compatibility to Julia and then save the jl-file (second argument) to the specified path:\\

\verb|using MatlabCompat|

\verb|rossetta(janus.m, janus.jl)|\\

One can test the functionality of this code using the minimalistic \href{https://github.com/MatlabCompat/MatlabCompat.jl/blob/master/test/janus.jl}{janus.jl} program discussed below. Unless currently in development, all the functions provided by the \href{https://github.com/MatlabCompat/MatlabCompat.jl}{MatlabCompat.jl} package have \href{https://github.com/MatlabCompat/MatlabCompat.jl/tree/master/test}{respective test programs} that exercise all the main features of the package for automated testing. Functionality of the whole package is continuously automatically tested using \href{https://travis-ci.org/MatlabCompat/MatlabCompat.jl}{Travis Continuouse Integrations} (CI) service.
  
  
  